### 2021-04-06 site/04-to-do.Rmd

# TODO

## Week 3

- [x] Check human subject experiment and Monash permission materials
	- https://www.intranet.monash/researchadmin/start/ethics/human
- [x] Check Kaiwen's project & paper
- [ ] Check Gallery of graphs - unclear author

## Week 4

- [x] Build prototype of html webpage to collect data
- [x] Send data to google sheet
- [x] Check summer-vis-inf 
- [x] Read examples sent by Di
- [ ] Build PhD repo
    - [x] meetings
    - [ ] paper
    - [x] TODO 
- [ ] Check Susan Vanderplas's website
    - [ ] paper
    - [ ] talks
    - [ ] posts

## Week 5

- [x] modify the webpage to be able to select multiple plots
- [x] Attempt to generate data from one assumed model
- [ ] draft Human Ethics Application Form

## Week 6

- [ ] Do the literature review of previous design
- [ ] Draw a flow chart to illustrate the design