### 2021-12-06 experiment/index.Rmd
add exp book

--- 
title: "Visual Inference for Diagnostic of Nonlinearity and Heteroskedasticy in Linear Regression"
author: "Patrick Li"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
# url: your book url like https://bookdown.org/yihui/bookdown
# cover-image: path to the social sharing image like images/cover.jpg
description: |
  .
biblio-style: apalike
csl: chicago-fullnote-bibliography.csl
---

# About

This book is about

## Structure of the book

In this book, will be covered ...