### 2021-12-06 experiment/05-data-quality.Rmd
add exp book

# (PART\*) Analysis of the experiment data {-}

# Data quality

The quality of the experiment data mainly depends on the respondent engagement. We could measure this by several variables:

1. response time
2. selections
3. Detection rate

Identifications for potential low quality responses:

  a. fail to identify an easy lineup because of insufficient response time
  b. significantly low detection rate because of insufficient response time throughout the study
  c. the detection rate adjusted by the number of selections significantly drop
  d.

## Study 1

## Study 2