### 2022-05-24 paper_comparison/explore/check_patterns.Rmd
explore patterns

---
title: "pattern_explore"
output: 
  html_document:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(visage)
library(tidyverse)
```

```{r fig.height=34, fig.width=45, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.25)
ori_dat <- mod$gen(1000, computed = list(e = rep(0, 1000)))
i <- 0
plot_list <- list()

for (a in seq(-1.5, 1.5, 0.25)) {
  for (b in seq(-1.5, 1.5, 0.25)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(1000, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE)} +
      ggtitle(glue::glue("a:{a}, b:{b}, a/|b|:{a/abs(b)}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 13)
```

There are 8 basic shapes, (1.5, 1.5), (1.5, -1.5), (1.5, 0), (-1.5, 0), (0, 1.5), (0, -1.5), (-1.5, 1.5), (-1.5, -1.5). Others are transitions between these five shapes.

```{r}
plot_list <- list()

mod$set_prm(c("a", "b"), c(-1.5, -1.5)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}")) -> plot_list[[1]]

mod$set_prm(c("a", "b"), c(-1.5, 0)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}")) -> plot_list[[2]]

mod$set_prm(c("a", "b"), c(-1.5, 1.5)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}")) -> plot_list[[3]]

mod$set_prm(c("a", "b"), c(0, -1.5)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}")) -> plot_list[[4]]

mod$set_prm(c("a", "b"), c(0, 1.5)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}")) -> plot_list[[6]]

mod$set_prm(c("a", "b"), c(1.5, -1.5)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}")) -> plot_list[[7]]

mod$set_prm(c("a", "b"), c(1.5, 0)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}")) -> plot_list[[8]]

mod$set_prm(c("a", "b"), c(1.5, 1.5)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}")) -> plot_list[[9]]

plot_list[[5]] <- patchwork::plot_spacer()

patchwork::wrap_plots(plot_list, ncol = 3)
```

These five are based on three patterns. Others are products of 180 degrees rotation or flip.

```{r fig.height=24, fig.width=15}
mod <- quartic_model(x = rand_uniform(-1, 1), sigma = 0.25)
ori_dat <- mod$gen(1000)
i <- 0
plot_list <- list()

for (c in seq(-20, 20, 5)) {
  i <- i + 1
  mod$set_prm(c("a", "b", "c"), c(5, 0, c)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}, c:{mod$prm$c}")) -> plot_list[[i]]

  i <- i + 1
  mod$set_prm(c("a", "b"), c(-5, 0)) %>%
    {.$gen(1000, computed = select(ori_dat, -y))} %>%
    {mod$plot(.)} +
    ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}, c:{mod$prm$c}")) -> plot_list[[i]]

  i <- i + 1
  mod$set_prm(c("a", "b"), c(-5, -5)) %>%
    {.$gen(1000, computed = select(ori_dat, -y))} %>%
    {mod$plot(.)} +
    ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}, c:{mod$prm$c}")) -> plot_list[[i]]

  i <- i + 1
  mod$set_prm(c("a", "b"), c(5, 5)) %>%
    {.$gen(1000, computed = select(ori_dat, -y))} %>%
    {mod$plot(.)} +
    ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}, c:{mod$prm$c}")) -> plot_list[[i]]

  i <- i + 1
  mod$set_prm(c("a", "b"), c(0, 5)) %>%
    {.$gen(1000, computed = select(ori_dat, -y))} %>%
    {mod$plot(.)} +
    ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}, c:{mod$prm$c}")) -> plot_list[[i]]
}

patchwork::wrap_plots(plot_list, ncol = 5)
```

The change of c brings two more shapes. (5,0,-5) and (-5,0,5). Other new shapes are quite similar to the previous. I don't think they are fundamentally different.

```{r}
mod$set_prm(c("a", "b", "c"), c(5, 0, -5)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}, c:{mod$prm$c}"))
```

```{r}
mod$set_prm(c("a", "b", "c"), c(-5, 0, 5)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}, c:{mod$prm$c}"))
```

Number count:

Consider using the simple cubic linear model.

$$y = 1 + x + ax^2 + bx^3$$