### 2022-06-13 paper_comparison/explore/check_different_sigma.Rmd
upload check model

---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(visage)
library(tidyverse)
```

# N: 50

Sigma: 0.125

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.125)
ori_dat <- mod$gen(50)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(50, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 0.25

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.25)
ori_dat <- mod$gen(50)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(50, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 0.5

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.5)
ori_dat <- mod$gen(50)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(50, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 1

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 1)
ori_dat <- mod$gen(50)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(50, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

# N: 100

Sigma: 0.125

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.125)
ori_dat <- mod$gen(100)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(100, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 0.25

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.25)
ori_dat <- mod$gen(100)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(100, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 0.5

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.5)
ori_dat <- mod$gen(100)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(100, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 1

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 1)
ori_dat <- mod$gen(100)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(100, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

# N: 300

Sigma: 0.125

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.125)
ori_dat <- mod$gen(300)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(300, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 0.25

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.25)
ori_dat <- mod$gen(300)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(300, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 0.5

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.5)
ori_dat <- mod$gen(300)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(300, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 1

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 1)
ori_dat <- mod$gen(300)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(300, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

# New design

# N: 50

Sigma: 1

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 1)
ori_dat <- mod$gen(50)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(50, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 2

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 2)
ori_dat <- mod$gen(50)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(50, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 5

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 5)
ori_dat <- mod$gen(50)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(50, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 10

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 10)
ori_dat <- mod$gen(50)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(50, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

# N: 100

Sigma: 1

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 1)
ori_dat <- mod$gen(100)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(100, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 2

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 2)
ori_dat <- mod$gen(100)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(100, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 5

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 5)
ori_dat <- mod$gen(100)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(100, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 10

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 10)
ori_dat <- mod$gen(100)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(100, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

# N: 300

Sigma: 1

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 1)
ori_dat <- mod$gen(300)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(300, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 2

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 2)
ori_dat <- mod$gen(300)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(300, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 5

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 5)
ori_dat <- mod$gen(300)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(300, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 10

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 10)
ori_dat <- mod$gen(300)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(300, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```