### 2023-02-06 paper_comparison/paper_comparison/paper_comparison.tex
add 1000 boot smooth lines for uniform (5)

% interactcadsample.tex
% v1.03 - April 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage{subfigure}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

\usepackage{natbib}% Citation support using natbib.sty
\bibpunct[, ]{(}{)}{;}{a}{}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{lscape}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\usepackage{xcolor}
\definecolor{orange-red}{rgb}{1.0, 0.27, 0.0}
\newcommand{\svp}[1]{{\textcolor{orange-red}{#1}}}
\newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

\articletype{ARTICLE TEMPLATE}

\title{Why is it that statistical tests for residuals are not widely
used? An explanation using visual inference.}

\author{\name{Weihao Li$^{a}$, Dianne Cook$^{a}$, Emi
Tanaka$^{a}$, Susan VanderPlas$^{b}$}
\affil{$^{a}$Department of Econometrics and Business Statistics, Monash
University, Clayton, VIC, Australia; $^{b}$Department of Statistics,
University of Nebraska, Lincoln, Nebraska, USA}
}

\thanks{CONTACT Weihao
Li. Email: \href{mailto:weihao.li@monash.edu}{\nolinkurl{weihao.li@monash.edu}}, Dianne
Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}, Emi
Tanaka. Email: \href{mailto:emi.tanaka@monash.edu}{\nolinkurl{emi.tanaka@monash.edu}}, Susan
VanderPlas. Email: \href{mailto:susan.vanderplas@unl.edu}{\nolinkurl{susan.vanderplas@unl.edu}}}

\maketitle

\begin{abstract}
Abstract to fill.
\end{abstract}

\begin{keywords}
data visualization; visual inference; hypothesis testing; residual
plots;
\end{keywords}

Title (jokes: finding the pattern in residual tests) (why using
visulization is important in regression diagnostics? an explanation
using visual inference) (a picture is worth a thousand word) (a plot is
worth a thousand test: )

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\begin{quote}
\emph{``Since all models are wrong the scientist must be alert to what
is importantly wrong.''} \citep{box1976science}
\end{quote}

Diagnosing a model is the key to determining whether there is anything
importantly wrong. In linear regression analysis, residuals are
typically examined for model diagnostics. Residuals summarise what is
not captured by the model, and thus provide the capacity to identify
what might be wrong.

We can assess residuals in multiple ways. Residuals might be plotted, as
a histogram or quantile-quantile plot to examine the distribution. Using
the classical normal linear regression model as an example, if the
distribution is symmetric and unimodal, we consider it well-behaved. But
if the distribution is skewed, bimodal, multimodal, or contains
outliers, there is cause for concern. One could also inspect the
distribution by conducting a goodness of fit test, such as the
Shapiro-Wilk Normality test \citep{shapiro1965analysis}.

More typically, residuals will be plotted, as a scatter plot against the
predicted values and each of the explanatory variables to scrutinize
their relationships. If there are any visually discoverable patterns,
the model is potentially misspecified. In general, one looks for
noticeable departures from the model like non-linear dependency or
heteroskedasticity. However, correctly judging a residual plot where no
pattern exists is a painstakingly difficult task for humans (?citation).
It is especially common, particularly among new data analysts, to report
patterns when an experienced one might quickly conclude that there are
none. It is also possible to conduct hypothesis tests for non-linear
dependence \citep{ramsey_tests_1969}, and use a Breusch-Pagan test
\citep{breusch_simple_1979} for heteroskedasticity.

Abundance of literature describe appropriate diagnostic methods for
linear regression: \citet{draper1998applied},
\citet{montgomery1982introduction}, \citet{belsley_regression_1980},
\citet{cook_applied_1999} and \citet{cook1982residuals}. The diligent
reader of these sage writings will also notice sentences that express
sentiments like \emph{based on their experience, statistical tests are
not widely used in regression diagnostics since the same or even larger
amount of information can be provided by diagnostic plots than the
corresponding tests in most empirical studies.} A common guidance by
experts is that optimal method for diagnosing model fits is by plotting
the data.

The persistence of this advice to check the plots is curious, and
investigating why this might be common advice is the subject of this
paper. The paper is structured as follows. The next background section
describes the types of departures that one expects to detect, and
outlines a formal statistical process for reading residual plots, called
visual inference. Section \ref{experimental-design} details the
experimental design to compare the decision made by formal hypothesis
testing, and how humans would read diagnostic plots. The results are
reported in Section \ref{results}. We finish with a discussion on future
work, in particular how the responsibility for residual plot reading
might be relieved.

\hypertarget{background}{%
\section{Background}\label{background}}

\hypertarget{departures-from-good-residual-plots}{%
\subsection{Departures from good residual
plots}\label{departures-from-good-residual-plots}}

(change it to facet)

(remove backticks for some plots)

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/residual-plot-common-departures-1}

}

\caption{Example residual vs fitted plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0. The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted plot.}\label{fig:residual-plot-common-departures}
\end{figure}

Graphical summaries in which residuals are plotted against fitted values
or other functions of the regressors that are approximately orthogonal
to residuals are referred to as standard residual plots in
\citet{cook1982residuals}. The first panel of Figure
\ref{fig:residual-plot-common-departures} shows an ideal residual plot
with residuals evenly distributed at both sides of the horizontal zero
line, with no noticeable patterns.

There are various types of departures from an ideal residual plot.
Non-linearity, heterskedasticity and non-normality are perhaps the three
mostly checked departures.

Non-linearity is a type of model misspecification caused by failing to
include higher order terms of the regressors in the regression equation.
Any non-linear functional form of residuals on fitted values in the
residual plot could be indicative of non-linearity. An example residual
plot containing visual pattern of non-linearity is given at the second
panel of Figure \ref{fig:residual-plot-common-departures}. One can
clearly observe the ``S-shape'' from the residual plot as the cubic term
is not captured by the misspecified model.

Heteroskedasticity refers to the presence of nonconstant error variance
in a regression model. It is mostly due to the strict but false
assumptions on the variance-covariance matrix of the error term. The
usual pattern of heteroskedasticity on a residual plot is the
inconsistent spread of the residuals across the horizontal axis.
Visually, it sometimes results in the so-called ``butterfly'' shape as
shown in the third panel of Figure
\ref{fig:residual-plot-common-departures}, or the ``left-triangle'' and
``right-triangle'' shape where the smallest variance occurs at one side
of the horizontal axis.

Compared to non-linearity and heteroskedasticity, non-normality is
usually harder to detect from a residual plot since a scatter plot do
not readily reveal the marginal distribution. A favourable graphical
summary for this task is the quantile-quantile plot. As we mainly
discuss residual plots, non-normality will not be the focus of this
paper. For a consistent comparison, the residual plot of this departure
is still presented in the fourth panel of Figure
\ref{fig:residual-plot-common-departures}. When the number residuals
below and above the horizontal axis are uneven across the local regions
along the \(x\)-axis, we expect that the normality assumption is
violated. For example, given a skewed error distribution, there will be
fewer data points and more outliers on one side of the horizontal axis
as shown in the fourth panel of Figure
\ref{fig:residual-plot-common-departures}.

\hypertarget{conventionally-testing-for-departures}{%
\subsection{Conventionally testing for
departures}\label{conventionally-testing-for-departures}}

Other than checking diagnostic plots, analysts may perform formal
hypothesis testing for detecting model defects. Depending on the
alternative hypothesis that is focused on, a variety of tests can be
applied. For example, the presence of heteroskedasticity can usually be
tested by applying the White test
\citep{white_heteroskedasticity-consistent_1980} or the Breusch-Pagan
test \citep{breusch_simple_1979}, which are both derived from the
Lagrange multiplier test \citep{silvey1959lagrangian} principle that
relies on the asymptotic properties of the null distribution. For
testing non-linearity, one may apply the F-test as a model structural
test to examine the significance of specific polynomial and non-linear
forms of the regressors, or the significance of proxy variables as in
the Ramsey Regression Equation Specification Error Test (RESET)
\citep{ramsey_tests_1969}. The Shapiro-Wilk test
\citep{shapiro1965analysis} is the most widely used test of
non-normality included by many of the statistical softwares. The
Jarque--Bera test \citep{jarque1980efficient} is also used to directly
checks if the sample skewness and kurtosis match a normal distribution.

Example residual plots given in Figure
\ref{fig:residual-plot-common-departures} are examined by the
corresponding RESET test, Breusch-Pagan test and Shapiro-Wilk test as
shown in Table \ref{tab:example-residual-plot-table}. In the example,
the Breusch-Pagan test and the Shapiro--Wilk test both reject the null
hypothesis \(H_0\) for departures that they do not intend to examine. As
discussed in \citet{cook1982residuals}, most residual-based tests for a
particular type of departure from model assumptions are also sensitive
to other types of departures. It is likely \(H_0\) is correctly rejected
but for the wrong reason, a phenomenon known as the ``Type III error''.
Additionally, outliers will often incorrectly trigger the rejection of
\(H_0\) despite when majority of the residuals are well-behaved
\citep{cook_applied_1999}. Furthermore, with a sufficiently large sample
size, residual-based tests may reject \(H_0\) due to a slight departure
that is of little practical significance. These can be largely avoided
in diagnostic plots as experienced analysts can evaluate the
acceptability of assumptions flexibly, even in the presence of outliers
and slight departures.

\begin{table}

\caption{\label{tab:example-residual-plot-table}Statistical significance testing for departures from good residuals for plots in Figure \ref{fig:residual-plot-common-departures}. Shown are the $p$-values calculated for the RESET, the Breusch-Pagan and the Shapiro–Wilk tests. The good residual plot (A) is judged a good residual plot, as expected, by all tests. The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.}
\centering
\begin{tabular}[t]{llrrr}
\toprule
Plot & Departures & RESET & Breusch-Pagan & Shapiro–Wilk\\
\midrule
A & None & 0.779 & 0.133 & 0.728\\
B & Non-linearity & \em{0.000} & \em{0.000} & \em{0.039}\\
C & Heteroskedasticity & 0.658 & \em{0.000} & \em{0.000}\\
D & Non-normality & 0.863 & 0.736 & \em{0.000}\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{visual-test-procedure-based-on-lineups}{%
\subsection{Visual test procedure based on
lineups}\label{visual-test-procedure-based-on-lineups}}

\hypertarget{lineup-protocol}{%
\subsubsection{Lineup protocol}\label{lineup-protocol}}

One may argue that reading diagnostic plots is to some extent subjective
and indecisive compared to those rigorous statistical procedures as it
relies on graphical perception - human ability to interpret and decode
the information embedded in graph \citep{cleveland_graphical_1984}.
Further, the degree of the presence of the visual features typically can
not be measured quantitatively and objectively, which may lead to over
or under-interpretations of the data. For instance, people
over-interpret the separation between gene groups in a two-dimensional
projection from a linear discriminant analysis when in fact there are no
differences in the expression levels between the gene groups and
separation is not an uncommon occurrence
\citep{roy_chowdhury_using_2015}.

Visual inference was first introduced in a 1999 Joint Statistical
Meetings (JSM) talk with the title ``Inference for Data Visualization''
by \citet{buja_inference_1999} as an idea to address the issue of valid
inference for visual discoveries of data plots. Later,
\citet{buja_statistical_2009} proposed the lineup protocol as a visual
test inspired by the ``police lineup'' or ``identity parade'' which is
the act of asking the eyewitness to identify criminal suspect from a
group of irrelevant people. The protocol consists of \(m\) randomly
placed plots, where one plot is the data plot, and the remaining
\(m - 1\) null plots have the identical graphical procedure except the
data has been replaced with data consistent with \(H_0\). Then, an
observer who have not seen the data plot will be asked to point out the
most different plot from the lineup. Under \(H_0\), it is expected that
the data plot would have no distinguishable difference from the null
plots, and the probability that the observer correctly picks the data
plot is \(1/m\). If one rejects \(H_0\) as the observer correctly picks
the data plot, then the Type I error of this test is \(1/m\).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/first-example-lineup-1}

}

\caption{Visual testing is conducted using a lineup, as in the example here. The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals are simulated from a standard error model. Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot. A small $p$-value would result from a substantial number selecting plot $2^2 + 2$.}\label{fig:first-example-lineup}
\end{figure}

Figure \ref{fig:first-example-lineup} is an example of a lineup
protocol. If the data plot at position \(2^2 + 2\) is identifiable, then
it is evidence for the rejection of \(H_0\) that the regression model is
correctly specified. In fact, the actual residual plot is obtained from
a misspecified regression model with non-linearity defect.

The effectiveness of lineup protocol for regression analysis is
validated by \citet{majumder_validation_2013} under relatively simple
settings with up to two regressors. Their results suggest that visual
tests are capable of testing the significance of a single regressor with
a similar power as a t-test, though they express that in general it is
unnecessary to use visual inference if there exists a conventional test,
and they do not expect the visual test to perform equally well as the
conventional test. (contradict to our claim? expand?) In their third
experiment, where there is not a conventional test, visual test
outperforms the conventional test for a large margin. This is
encouraging, as it promotes the use of visual inference in situations
where there are no existing statistical testing procedures. Visual
inference have also been integrated into diagnostic of hierarchical
linear models by \citet{loy2013diagnostic}, \citet{loy2014hlmdiag} and
\citet{loy2015you}. They use lineup protocols to judge the assumption of
linearity, normality and constant error variance for both the level-1
and level-2 residuals. (expand?)

\hypertarget{sampling-from-the-null-distribution}{%
\subsubsection{Sampling from the null
distribution}\label{sampling-from-the-null-distribution}}

Data used in the \(m - 1\) null plots needs to be simulated. In
regression diagnostics, sampling data consistent with \(H_0\) is
equivalent to sampling data from the assumed model. As
\citet{buja_statistical_2009} suggested, \(H_0\) is usually a composite
hypothesis controlled by nuisance parameters. Since regression models
can have various forms, there is no general solution to this problem,
but it sometimes can be reduced to a so called ``reference
distribution'' by applying one of the three methods: (i) sampling from a
conditional distribution given a minimal sufficient statistic under
\(H_0\), (ii) parametric bootstrap sampling with nuisance parameters
estimated under \(H_0\), and (iii) Bayesian posterior predictive
sampling. The conditional distribution given a minimal sufficient
statistic is the best justified reference distribution among the three
\citep{buja_statistical_2009}. Essentially, null residuals can be
simulated by regressing \(N\) i.i.d standard normal random draws on the
regressors, then rescaling it by the ratio of residual sum of square in
two regressions.

\hypertarget{calculating-p-values-for-the-visual-test}{%
\subsubsection{\texorpdfstring{Calculating \(p\)-values for the visual
test}{Calculating p-values for the visual test}}\label{calculating-p-values-for-the-visual-test}}

In hypothesis testing, a \(p\)-value is defined as the probability of
observing test results as least as extreme as the observed result given
\(H_0\) is true. Within the context of visual inference, by involving
\(k\) independent observers, the \(p\)-value can be interpreted as the
probability of having as many or more subjects detect the data plot than
the observed result.

Let \(X_j = \{0,1\}\) be a Bernoulli random variable denoting whether
subject \(j\) correctly detecting the data plot, and
\(X = \sum_{j=1}^{K}X_j\) be the number of observers correctly picking
the data plot. Then, by imposing a relatively strong assumption on the
visual test that all \(K\) evaluations are fully independent, under
\(H_0\), \(X \sim \mathrm{Binom}_{K,1/m}\). Therefore, the \(p\)-value
of a lineup of size \(m\) evaluated by \(K\) observer is given as
\(P(X \geq x) = 1 - F(x) + f(x)\), where \(F(.)\) is the cumulative
distribution function, \(f(.)\) is the probability mass function and
\(x\) is the realization of number of observers correctly picking the
data plot \citep{majumder_validation_2013}.

As pointed out by \citet{vanderplas2021statistical}, this basic binomial
model doesn't take into account the possible dependencies in the visual
test due to repeated evaluations of the same lineup. And it is
inapplicable to visual test where subjects are asked to select one or
more ``most different'' plots from the lineup.
\citet{vanderplas2021statistical} summarises three common scenarios in
visual inference: (1) \(K\) different lineups are shown to \(K\)
subjects, (2) \(K\) lineups with different null plots but the same data
plot are shown to \(K\) subjects, and (3) the same lineup is shown to
\(K\) subjects. Out of these three scenarios, Scenario 3 is the most
common in previous studies as it puts the least constraints on the
experiment design. For Scenario 3, \citet{vanderplas2021statistical}
models the probability of a plot \(i\) being selected from a lineup as
\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
\(i=1,...,m\) and \(\alpha > 0\). The number of times plot \(i\) being
selected in \(K\) evaluations is denoted as \(c_i\). In case subject
\(j\) makes multiple selections, \(1/s_j\) will be added to \(c_i\)
instead of one, where \(s_j\) is the number of plots subject \(j\)
selected for \(j=1,...K\). This ensures \(\sum_{i}c_i=K\). Since we are
only interested in the selections of the data plot \(i\), the marginal
model can be simplified to a beta-binomial model and thus the visual
p-value is given as

\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+
\end{equation}

\noindent where \(B(.)\) is the beta function defined as

\begin{equation} \label{eq:betafunction}
B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0. 
\end{equation}

Note that Equation \ref{eq:pvalue-beta-binomial} given in
\citet{vanderplas2021statistical} only works with non-negative integer
\(c_i\). We extend the equation to non-negative real number \(c_i\) by
applying a linear approximation

\begin{equation} \label{eq:pvalue-beta-binomial-approx}
P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+,
\end{equation}

where \(P(C \geq \lceil c_i \rceil)\) is calculated using Equation
\ref{eq:pvalue-beta-binomial} and \(P(C = \lfloor c_i \rfloor)\) is
calculated by

\begin{equation} \label{eq:pmf-beta-binomial}
P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
\end{equation}

Besides, the parameter \(\alpha\) used in Equation
\ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually
unknown and hence needs to be estimated from the survey data. For low
values of \(\alpha\), only a few plots are attractive to the observers
and tend to be selected. For higher values of \(\alpha\), the
distribution of the probability of each plot being selected is more
even. \citet{vanderplas2021statistical} defines that a plot is
\(c\)-interesting if \(c\) or more participants select the plot as the
most different. Given the definition, The expected number of plots
selected at least \(c\) times, \(E[Z_c]\), is calculated as

\begin{equation} \label{eq:c-interesting-expectation}
E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}

With Equation \ref{eq:c-interesting-expectation}, \(\alpha\) can be
estimated using maximum likelihood estimation. But for precise estimate
of \(\alpha\), additional responses to Rorschach lineups, which is a
type of lineup that consists of plots constructed from the same null
data generating mechanism, are required.

\hypertarget{power-of-a-visual-test}{%
\subsubsection{Power of a visual test}\label{power-of-a-visual-test}}

The power of a model misspecification test is the probability that the
null hypothesis is rejected given the regression model is misspecified.
It is an important indicator when one is concerned about whether model
assumptions have been violated. Although in practice, one might be more
interested in knowing how much the residuals deviate from the model
assumptions, and whether this deviation is of practical significance.

As discussed in \citet{majumder_validation_2013}, the power of a visual
test may depend on the ability of the particular subject, as the skill
of the individual may affect the number of observers who identify the
data plot from the lineup. Previously, it is addressed by modelling the
probability of a subject \(j\) correctly picking the data plot from a
lineup \(l\) using a mixed-effect logistic regression, with subjects
treated as random effects \citep{majumder_validation_2013}. However, in
the multiple selections scenario, having this probability is not
sufficient to determine the power of a visual test because it does not
provide information about the number of selections made by the subject
for the calculation of the p-value.

Instead, we directly estimate the probability of a lineup being rejected
by assuming that individual skill has negligible effect on the variation
of the power. The assumption is to simplify the model structure, thereby
obviate costly large-scale experiments to estimate complex covariance
matrices. The model is a logistic regression with the natural logarithm
of the effect as the only regressor formulated as

\begin{equation} \label{eq:logistic-regression-1-1}
Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log(E)),
\end{equation}

\noindent where \(\Lambda(.)\) is the standard logistic function given
as \(\Lambda(z) = exp(z)/(1+exp(z))\).

Effect \(E\) is derived from the Kullback-Leibler divergence (see
Appendix for the details) formulated as

\small

\begin{gather}  \label{eq:effect-size}
E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V})|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V})^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V})^{-1}\boldsymbol{\mu}_z\right),\\
\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a,\\
\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a^T\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a^T,
\end{gather}

\normalsize

\noindent where \(diag(.)\) is the diagonal matrix constructed from the
diagonal elements of a matrix,
\(\boldsymbol{X}_a = (\boldsymbol{1}, \boldsymbol{X})\) is the matrix of
regressors including the intercept used in the regression equation,
\(\sigma^2\boldsymbol{I}\) is the assumed variance of the error term
when \(H_0\) is true, \(\boldsymbol{V}\) is the actual variance of the
error term, and
\(\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z\)
is the expected values of the residuals with \(\boldsymbol{Z}\) be any
variables leave out by the model and \(\boldsymbol{\beta}_z\) be the
corresponding coefficients.

To study various factors contributing to the power of the visual test,
the same logistic regression model is fit on different subsets of the
collated data grouped by levels of factors. These include the
distribution of the regressor and the type of the simulation model
(modify this according to the result).

\hypertarget{experimental-design}{%
\section{Experimental design}\label{experimental-design}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  experiments -\textgreater{} data collection periods
\end{enumerate}

To investigate the difference between conventional hypothesis testing
and visual inference in the application of linear regression
diagnostics, an experiment is conducted in three data collection
periods. Two types of departures, namely non-linearity and
heteroskedasticity, are considered with the corresponding data
generating process being designed for data collection periods I and II.
The data collection period III is designed for collecting human
responses to null lineups such that the parameter \(\alpha\) in Equation
\ref{eq:pvalue-beta-binomial} can be estimated. Overall, we plan to
collect 7974 evaluations on 1152 unique lineups performed by 443
subjects throughout three data collection periods.

\hypertarget{simulating-departures-from-good-residuals}{%
\subsection{Simulating departures from good
residuals}\label{simulating-departures-from-good-residuals}}

\hypertarget{non-linearity}{%
\subsubsection{Non-linearity}\label{non-linearity}}

Data collection period I is designed to study the ability of human
subjects to detect the effect of a non-linear term \(\boldsymbol{z}\)
constructed using Hermite polynomials on random vector
\(\boldsymbol{x}\) formulated as

\begin{align} \label{eq:nonlinearity-model}
\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
\end{align}

\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
\(\boldsymbol{\varepsilon}\), \(\boldsymbol{x}_{raw}\),
\(\boldsymbol{z}_{raw}\) are vectors of size \(n\), \(He_{j}(.)\) is the
\(j\)th-order probabilist's Hermite polynomials,
\(\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\), and
\(g(\boldsymbol{x}, k)\) is a scaling function to enforce the support of
the random vector to be \([-k, k]^n\) defined as

\begin{equation} \label{eq:scaling-function}
g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x}))2k - k, \quad \text{for} \quad k > 0. 
\end{equation}

According to \citet{abramowitz1964handbook}, Hermite polynomials were
initially defined by \citet{de1820theorie}, but named after Hermite
\citep{hermite1864nouveau} because of the unrecognisable form of
Laplace's work. When simulating \(\boldsymbol{z}_{raw}\), function
\texttt{hermite} from the R package \texttt{mpoly} \citep{mpoly} is used
to generate Hermite polynomials.

The null regression model used to fit the realizations generated by the
above model is formulated as

\begin{equation} \label{eq:null-model}
\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
\end{equation}

\noindent where
\(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).

Since \(z = O(x^j)\), for \(j > 1\), \(z\) is a higher order term leaves
out by the null regression, which will lead to model misspecification.

Visual patterns of non-linearity are simulated using four different
orders of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)). (A
summary of the factors is given in Table \ref{tab:model-factor-table}.)
The values of \(j\) is chosen so that distinct shapes of non-linearity
are included in the residual plot. These include ``U'', ``S'', ``M'' and
``Triple-U'' shape as shown in Figure
\ref{fig:different-shape-of-herimite}. A greater value of \(j\) will
result in a curve with more turning points. It is expected that the
``U'' shape will be the easiest one to detect because complex shape
tends to be concealed by cluster of data points.

\begin{table}

\caption{\label{tab:model-factor-table}Levels of the factors used in data collection periods I, II, III.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{rr|rr|rcrr|rr|rcrr|rr|rcrr|rr|rcrr|rr|rcrr|rr|rc}
\toprule
\multicolumn{2}{c}{Non-linearity} & \multicolumn{2}{c}{Heteroskedasticity} & \multicolumn{2}{c}{Common} \\
\cmidrule(l{3pt}r{3pt}){1-2} \cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6}
\multicolumn{1}{c}{Poly Order ($j$)} & \multicolumn{1}{c}{SD ($\sigma$)} & \multicolumn{1}{c}{Shape ($a$)} & \multicolumn{1}{c}{Ratio ($b$)} & \multicolumn{1}{c}{Size ($n$)} & \multicolumn{1}{c}{Distribution of fitted values} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
2 & 0.25 & -1 & 0.25 & 50 & Uniform\\
3 & 1.00 & 0 & 1.00 & 100 & Normal\\
6 & 2.00 & 1 & 4.00 & 300 & Skewed\\
18 & 4.00 &  & 16.00 &  & Discrete uniform\\
 &  &  & 64.00 &  & \\
\bottomrule
\end{tabular}}
\end{table}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-herimite-1}

}

\caption{Polynomial forms generated for the residual plots used in data collection period I. The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.}\label{fig:different-shape-of-herimite}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-dist-1}

}

\caption{Variations in fitted values, that might affect perception of residual plots. Four different distributions are used.}\label{fig:different-dist}
\end{figure}

Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups
used in data collection period I. This lineup is produced by the
non-linearity model with \(j = 6\) and uniform fitted values. The data
plot location is \(2^3 - 4\). All five subjects correctly identify the
data plot from this lineup.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-n-1}

}

\caption{Number of observations $n$ have an impact on the strength of the signal. Three different values of $n$ are used in data collection periods I, II and III.}\label{fig:different-n}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-sigma-1}

}

\caption{For the non-lienearity model, the standard deviation of the error term $\sigma$ affects the strength of the signal. Four different values of $\sigma$ are used in the data collection period I.}\label{fig:different-sigma}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-poly-lineup-1}

}

\caption{One of the lineups containing non-linearity patterns used in data collection period I. Can you spot the most different plot? The data plot is positioned at $2^3 + 1$.}\label{fig:example-poly-lineup}
\end{figure}

\hypertarget{heteroskedasticity}{%
\subsubsection{Heteroskedasticity}\label{heteroskedasticity}}

Data collection period II is designed to study the ability of human
subjects to detect the appearance of a heteroskedasticity pattern under
a simple linear regression model setting:

\begin{align} \label{eq:heter-model}
\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), 
\end{align}

\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
\(\boldsymbol{\varepsilon}\) are vectors of size \(n\) and \(g(.)\) is
the scaling function defined in Equation \ref{eq:scaling-function}.

The null regression model used to fit the realizations generated by the
above model is formulated exactly the same as Equation
\ref{eq:null-model}.

For \(b \neq 0\), the variance-covariance matrix of the error term
\(\boldsymbol{\varepsilon}\) is correlated with the regressor
\(\boldsymbol{x}\), which will lead to the presence of
heteroskedasticity. Visual patterns of heteroskedasticity are simulated
using three different shapes (\(a\) = -1, 0, 1). (A summary of the
factors can be found in Table \ref{tab:model-factor-table}.)

Since \(supp(X) = [-1, 1]\), choosing \(a\) to be \(-1\), \(0\) and
\(1\) can generate ``left-triangle'', ``butterfly'' and
``right-triangle'' shape as displayed in Figure
\ref{fig:different-shape-of-heter}. The term \((2 - |a|)\) maintains the
magnitude of residuals across different values of \(a\).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-heter-1}

}

\caption{Heteroskedasticity forms used in data collection period II. Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle, butterfly and right-triangle shapes.}\label{fig:different-shape-of-heter}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-b-1}

}

\caption{Five different values of $b$ are used in data collection period II to control the strength of the signal.}\label{fig:different-b}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-heter-lineup-1}

}

\caption{One of the lineups containing heteroskedasticity pattern used in data collection period II. Can you spot the most different plot? The data plot is positioned at $3^3 - 3^2$}\label{fig:example-heter-lineup}
\end{figure}

An example lineup of this model used in data collection period II is
shown in Figure \ref{fig:example-heter-lineup} with \(a = -1\) and
\(X_{raw} \sim U(-1, 1)\). The data plot location is \(2^4 + 2\). Nine
out of 11 subjects correctly identify the data plot from this lineup.

\hypertarget{factors-common-to-both-data-collection-periods}{%
\subsubsection{Factors common to both data collection
periods}\label{factors-common-to-both-data-collection-periods}}

XXX DC will clean this up

Four distributions of fitted values are also considered by using
\(X_{raw}\) follows different distributions: (1) \(U(-1, 1)\), (2)
\(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\).

Different distributions of fitted values help enriching the pool of
visual patterns as illustrated in Figure \ref{fig:different-dist}. The
uniform and the normal distribution are symmetric and commonly assumed
in statistical models. The adjusted log-normal distribution provides
skewed density, while the discrete uniform distribution provides
discreteness in residual plot.

\hypertarget{experimental-setup}{%
\subsection{Experimental setup}\label{experimental-setup}}

\hypertarget{controlling-the-strength-of-the-signal}{%
\subsubsection{Controlling the strength of the
signal}\label{controlling-the-strength-of-the-signal}}

As summarised in Table \ref{tab:model-factor-table}, three additional
parameters \(n\), \(\sigma\) and \(b\) are used to control the strength
of the signal so that different difficulty levels of lineups are
generated, and therefore, the estimated power curve will be smooth and
continuous. Parameter \(\sigma \in \{0.5, 1, 2, 4\}\) and
\(b \in \{0.25, 1, 4, 16, 64\}\) are used in data collection periods I
and II respectively. Figure \ref{fig:different-sigma} and
\ref{fig:different-b} demonstrate the impact of these two parameters. A
large value of \(\sigma\) will increase the variation of the error of
the non-linearity model and decrease the visibility of the visual
pattern. The parameter \(b\) controls the standard deviation of the
error across the support of the regressor. Given \(x \neq a\), a larger
value of \(b\) will lead to a larger ratio of the variance at \(x\) to
the variance at \(x - a = 0\), making the visual pattern more obvious.

Three different sample sizes are used (n = 50, 100, 300) in all three
data collection periods. It can be observed from Figure
\ref{fig:different-n} that with fewer data points drawn in a residual
plot, the visual pattern is more difficult to be detected.

\hypertarget{subject-allocation}{%
\subsubsection{Subject allocation}\label{subject-allocation}}

Three replications are made for each of the parameter values shown in
Table \ref{tab:model-factor-table} resulting in
\((4 \times 4 \times 4 \times 3 + 4 \times 3 \times 5 \times 3) \times 3 = 1116\)
different lineups. In addition, each lineup is designed to be evaluated
by five different subjects. After attempting some pilot studies
internally in our research group, we decide to present a block of 20
lineups to every subject. And to ensure the quality of the survey data,
two lineups with obvious visual patterns are included as attention
checks. Thus, \(576 \times 5 / (20-2) = 160\) and
\(540 \times 5 / (20-2) = 150\) subjects are recruited to satisfy the
design of the data collection period I and II respectively.

As mentioned in Section \ref{calculating-p-values-for-the-visual-test},
\(\alpha\) used in Equation \ref{eq:pvalue-beta-binomial} needs to be
estimated using null lineups. Hence, three replications of 36 lineups
with all combinations of \(n\) and fitted value distributions are
included in data collection period III. In these lineups, the data of
the data plot is generated from a model with zero effect size, while the
data of the 19 null plots are generated using the same simulation method
discussed in Section \ref{sampling-from-the-null-distribution}. This
generation procedure differs from the canonical Rorschach lineup
procedure, which requires that all 20 plots are generated from the null
hypothesis. However, these lineups serve the same fundamental purpose:
to assess the number of visually interesting plots generated under the
null hypothesis.

To account for the fact that our simulation method for these lineups is
not the Rorschach procedure, we use the method suggested in
\citet{vanderplas2021statistical} for typical lineups containing a data
plot to estimate \(\alpha\). We have included a sensitivity analysis in
the Appendix to examine the impact of the variance of the \(\alpha\)
estimate on our findings.

All lineups consist of only null plots are planned to be evaluated by 20
subjects. However, presenting only these lineups to subjects are
considered to be bad practices as subjects will lose interest quickly.
Therefore, we plan to collect 6 more evaluations on the 279 lineups with
uniform fitted values, result in
\((36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = 133\)
subjects recruited for data collection period III.

\hypertarget{collecting-results}{%
\subsubsection{Collecting results}\label{collecting-results}}

Subjects for all three data collection periods are recruited from an
crowdsourcing platform called Prolific \citep{palan2018prolific}.
Prescreening procedure is applied during the recruitment, subjects are
required to be fluent in English, with \(98\%\) minimum approval rate
and 10 minimum submissions in other studies.

During the experiment, every subject is presented with a block of 20
lineups. A lineup consists of a randomly placed data plot and 19 null
plots, which are all residual plots drawn with raw residuals on the
y-axis and fitted values on the x-axis. An additional horizontal red
line is added at \(y = 0\) as a helping line.

The data of the data plot is simulated from one of two models described
in Section \ref{simulating-departures-from-good-residuals}, while the
data of the remaining 19 null plots are generated by the residual
rotation technique discussed in Section
\ref{sampling-from-the-null-distribution}.

In every lineup evaluation, the subject is asked to select one or more
plots that are most different from others, provide a reason for their
selections, and evaluate how different they think the selected plots are
from others. If there is no noticeable difference between plots in a
lineup, subjects are permitted to select zero plots without providing
the reason. No subject are shown the same lineup twice. Information
about preferred pronoun, age group, education, and previous experience
in visual experiments are also collected. A subject's submission is only
accepted if the data plot is identified for at least one attention
check. Data of rejected submissions are discarded automatically to
maintain the overall data quality.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{overview}{%
\subsection{Overview}\label{overview}}

There are 2880, 2700 and 1674 lineups evaluation made by 160, 150 and
133 subjects recruited for data collection periods I, II and III
respectively. In the total of 7974 lineup evaluations, 3744 use lineups
produced by the non-linearity model, and 4230 use lineups produced by
the heteroskedasticity model. Besides, there are 886 attention checks
and 720 evaluations on null lineups needed for the estimate of
\(\alpha\) not included in the analysis. The collated dataset is
provided in \texttt{vi\_survey} of the \texttt{visage} \texttt{R}
package.

In the following analysis, lineups with uniform fitted values will be
the focus. Visual patterns are more likely to be revealed under a
uniform distribution. Additionally, we have collected extra evaluations
on these lineups, which will result in more reliable analysis. Analysis
of lineups with other fitted value distributions can be found in Section
\ref{effect-of-other-fitted-value-distributions}.

\hypertarget{power-comparison-of-different-tests}{%
\subsection{Power comparison of different
tests}\label{power-comparison-of-different-tests}}

Figure \ref{fig:polypower} shows the estimated power of visual test on
lineups produced by the non-linearity model with uniform fitted values,
against the natural logarithm of the effect \(log_e(E)\), with a 5\%
significance level. At the bottom of the figure \ref{fig:polypower},
there are a sequence of example residual plots with increasing levels of
\(log_e(E)\). Readers can evaluate them from left to right and determine
at which level the departure from a good residual plot becomes
detectable.

As discussed in Section \ref{conventionally-testing-for-departures},
many conventionally tests are available for detecting residual
departures. Implementation-wise, the built-in R package \texttt{stats}
provides some commonly used residual-based tests, such as Shapiro-Wilk
test. A more comprehensive collection of regression diagnostics tests
can be found in the R package \texttt{lmtest} \citep{lmtest}. In terms
of heteroskedasticity diagnostics, the R package \texttt{skedastic}
\citep{skedastic} collects and implements 25 existing conventional tests
published since 1961.

To compare the power of visual test and conventional test, we pick RESET
test (\texttt{resettest}) and Breusch-Pagan test (\texttt{bptest}) from
the R package \texttt{lmtest}, and Shapiro-Wilk test
(\texttt{shapiro.test}) from the built-in R package \texttt{stats}.
Among them, RESET test is the only exact and appropriate test in this
scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are
approximate and inappropriate tests. Their estimated power is shown in
Figure \ref{fig:polypower}. To set up the RESET test, we include
different powers of fitted values as proxies. According to
\citet{ramsey_tests_1969}, there are no general rules for the power of
the fitted values needed by the RESET test, but it finds power up to
four is usually sufficient. Thus, we follow this guideline to conduct
the RESET test. For the Breusch-Pagan test, the choice of regressors in
the auxiliary regression is left to the user
\citep{breusch_simple_1979}. But as \citet{waldman1983note} suggested,
it is a good choice for the set of auxiliary regressors in the
Breusch-Pagan test be the same as the White test. Thus, we include both
\(\boldsymbol{x}\) and \(\boldsymbol{x}^2\) in the auxiliary regression.

Figure \ref{fig:heterpower} is similar to Figure \ref{fig:polypower},
but shows corresponding information on lineups produced by the
heteroskedasticity model. In this scenario, the visual test is compared
to an approximate test - Breusch-Pagan test, and two other inappropriate
tests - RESET test and Shapiro-Wilk test.

For non-linearity patterns, the power curve of RESET test climbs
aggressively from 15\% to 69\% as \(log_e(E)\) increases from 0 to 2,
while power of other tests respond inactively to the change of effect,
showing that RESET test is way more sensitive to the type of model
defects that being considered. Meanwhile, no noticeable visual features
can be spotted from the example residual plots.

In terms of heteroskedasticity patterns, the power of Breusch-Pagan test
is also almost always greater than the power of visual test. For
\(0 \leq log_e(E) \leq 2\), where the power curve of the visual test
remains at a low level, the Breusch-Pagan test still have a decent
amount of chance of rejecting \(H_0\). Similarly, the visual feature is
nearly unobservable from the example residual plots.

The power of visual test arises steadily as \(log_e(E)\) increases from
2 to 5 for both non-linearity patterns and heteroskedasticity patterns,
suggesting that the effect starts to make significant impact on the
degree of the presence of the designed visual features. This can also be
observed from the example residual plots that when \(log_e(E) = 2.5\), a
weak ``S-shape'' and a weak ``triangle'' shape are presented in Figure
\ref{fig:polypower} and Figure \ref{fig:heterpower} respectively. The
visual pattern becomes much clearer as \(log_e(E)\) increases. At
\(log_e(E) \approx 6\), the power reaches almost 100\%.

The power of all inappropriate tests except for RESET test shows
improvement as the effect increases but at a lower rate than the visual
test in both scenarios. This coincides the point made by
\citet{cook1982residuals} that residual-based tests for a specific type
of model defect may be sensitive to other types of model defects. The
power curve of RESET test remains at around 5\% in Figure
\ref{fig:heterpower} since there are no non-linear terms leave out in
the heteroskedasticity model and \(H_0\) of the test is always
satisfied.

Overall, the power comparison suggests that conventional tests differs
significantly from visual tests in two regression diagnostics scenarios
designed by us. Visual test have much higher tolerance of the residual
departures than the conventional test. Since fail to reject \(H_0\) in a
visual test usually means that there are no obvious visual discoveries
found in the residual plot, analysts and the general public as the
consumers of the output may not be convinced of the existence of
significant residual departures in spite of the rejection of \(H_0\)
given by the conventional test. Even if the rejection is accepted, the
model violation may be considered as impactless due to the fact that
they are not clearly visible. Besides, the sensitivity of the
conventional test could also distract and discourage analysts from
finding simple but good linear approximation to the data. The rejection
of \(H_0\) because of human acceptable and negligible residual
departures is not practically meaningful and useful. This may limit the
popularity of conventional tests in residual diagnostics among analysts.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/polypower-1}

}

\caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only). Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups. Surrounding lines of the visual test show the estimated power of 1000 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot. Where would you draw the line of too much non-linearity in the residuals? For the RESET test this is around log effect size 1.5, but for the visual test it is around 3.5.}\label{fig:polypower}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heterpower-1}

}

\caption{Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only). Main plot shows the power curves, with dots indicating human evaluations of lineups. Surrounding lines of the visual test show the estimated power of 1000 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot. Where would you draw the line of too much heteroskedasticity in the residuals? For the BP test this is around log effect size 1.5, but for the visual test it is around 3.}\label{fig:heterpower}
\end{figure}

\hypertarget{comparison-of-test-decisions-based-on-p-values}{%
\subsection{\texorpdfstring{Comparison of test decisions based on
\(p\)-values}{Comparison of test decisions based on p-values}}\label{comparison-of-test-decisions-based-on-p-values}}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/p-value-comparison-1}

}

\caption{Rejection rate ($p$-value $<0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot. The visual test rejects less frequently than the conventional test. We would generally expect that the visual test would only reject when the conventional test does. Surprisingly, there is one lineup in the heteroskedasticity study where this is not the case. }\label{fig:p-value-comparison}
\end{figure}

The power comparison illustrates that appropriate conventional tests
will reject \(H_0\) more aggressively than visual tests. In this
section, we explore how often they agree with each other by
investigating the rejections for the two model designs based on
\(p\)-values for each lineup.

Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the
rejection rate of visual tests and conventional tests for both
non-linearity patterns and heteroskedasticity patterns.

For lineups containing non-linearity patterns, conventional tests reject
69\% and visual tests reject 32\% of the time. Of the lineups rejected
by the conventional test, 46\% are rejected by the visual test, that is,
approximately half as many as the conventional test. There are no
lineups that are rejected by the visual test but not by the conventional
test.

In terms of lineups containing heteroskedasticity patterns, 76\% are
rejected by conventional tests, while 56\% are rejected by visual tests.
When the conventional test rejects a lineup, there is a great chance
(73\%) the visual test will also reject it.

Surprisingly, the visual test rejects 1 of the 33 (3\%) of lineups where
the conventional test does not reject. Figure \ref{fig:heter-example}
shows this lineup. The data plot in position seventeen displays a
relatively strong heteroskedasticity pattern, and has a strong effect
size (\(log_e E=4.02\)). This is reflected by the visual test
\(p\text{-value} = 0.026\), but the Breusch-Pagan test
\(p\text{-value} = 0.056\), slightly above the significance cutoff of
\(0.05\). This lineup was evaluated by 11 subjects, it has experimental
factors \(n=50\) (small sample size), \(\sigma=1\), and a uniform
distribution for the fitted values. It must be the small sample size
that may have resulted in the lack of detection.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heter-example-1}

}

\caption{The single heteroskedasticity lineup that is rejected by the visual test but not by the BP test. The data plot at panel 17 contains a "Butterfly" shape. It has effect size $ = 4.02$, somewhat surprising that it is not detected by the BP test.}\label{fig:heter-example}
\end{figure}

\hypertarget{effect-of-shape-of-non-linearity}{%
\subsection{Effect of shape of
non-linearity}\label{effect-of-shape-of-non-linearity}}

A primary factor contributes to the non-linearity model is the shape of
the non-linearity. According to Figure \ref{fig:poly-power-uniform-j},
conventional tests have almost identical power in testing the ``U'', the
``S'' and the ``M'' shape, except for testing the ``Triple-U'' shape,
which have significant lower power. To understand why this is, one needs
to return to the way the RESET test is applied. It requires a parameter
indicating degree of fitted values to test for, and the recommendation
is to generically use four (\citet{ramsey_tests_1969}). However, the
``Triple-U'' shape constructed from the Hermite polynomials use power up
to 18. If the RESET test had been applied using a higher power no less
than six, the power curve of ``Triple-U'' shape will have little
difference to other shapes. The recommendation of the polynomial power
for the RESET should be revised, perhaps. This illustrates the
sensitivity of conventional testing to the parameters, and it also
points to a limitation that one needs to know the data structure in
order to set the parameters for the test.

For visual tests, we expect the ``U'' shape will be the easiest one to
be detected by subjects followed by the ``S'', ``M'' and ``Triple-U''
shape. From Figure \ref{fig:poly-power-uniform-j}, it can be observed
that the power curves are mostly aligned with our expectation, except
for the ``M'' shape, which is as easy to be detected as the ``U'' shape.
This implies the readability of the shape do not strictly follow the
degree of the polynomials.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/poly-power-uniform-j-1}

}

\caption{Power of conventional tests and visual tests on lineups containing four different non-linearity shapes. Power curves with higher order of non-linearity are drawn with deeper colours. The default RESET tests under-perform significantly in detecting "Tripe-U" shape. To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test. But this means the order needs to be known prior to testing.}\label{fig:poly-power-uniform-j}
\end{figure}

\hypertarget{effect-of-shape-of-heteroskedasticity}{%
\subsection{Effect of shape of
heteroskedasticity}\label{effect-of-shape-of-heteroskedasticity}}

We also investigate the impact of different heteroskedasticity shapes on
power of conventional tests and visual tests. In theory, the
``Left-triangle'' and the ``Right-triangle'' shapes are functionally
identical from the point of view of a Breusch-Pagan test. As shown in
Figure \ref{fig:heter-power-uniform-a}, this is indeed the case where
little difference between the power curves can be perceived. Similarly,
visual tests will have the same power of detecting these two shapes if
they are equally likely to be identified. However, it can be observed
from Figure \ref{fig:heter-power-uniform-a} that the power curve for the
``Left-triangle'' shape is constantly higher than the one for the
``Right-triangle'' shape, indicating a potential favour of orientation
by human, which is worth to be explored in future studies.

The chance of detecting the ``Butterfly'' shape is roughly the same as
the chance of detecting the ``Left-triangle'' shape in both tests. The
only noticeable difference from Figure \ref{fig:heter-power-uniform-a}
is the Breusch-Pagan test has slightly higher power for detecting this
shape compared to other two shapes when the effect size is low.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heter-power-uniform-a-1}

}

\caption{Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$.}\label{fig:heter-power-uniform-a}
\end{figure}

\hypertarget{effect-of-fitted-value-distributions}{%
\subsection{Effect of fitted value
distributions}\label{effect-of-fitted-value-distributions}}

The prediction in a regression model is \(E(Y|X)\), that is, it is
conditional on observed values of the predictors. The distribution of
\(X\), or consequently \(\hat{Y}\), may however affect the ability to
read any patterns in the residuals. The four distributions of fitted
values, used in this experiment, were designed to examine this.

Figure \ref{fig:different-x-dist-poly-power} illustrates the change of
power of visual tests and conventional tests for different fitted
distributions used in the non-linearity model and the heteroskedasticity
model. We focus on \(log(E) > 2\) where the visual patterns start to
become recognisable.

For non-linearity patterns, we do not observe significant difference in
power of visual test between distributions, except for the discrete
uniform distribution. It could be due to the fact that humans have
difficulties in recognising visual patterns from discreteness, which
makes the shape disconnected and incomplete (ref?). In terms of
conventional tests, the discrete uniform distribution is instead the one
with the greatest power, followed by the lognormal distribution, the
uniform distribution and the normal distribution.

For heteroskedasticity patterns, the power of visual test on lineups
with uniform distribution has the greatest power. This is as expected
since other three distributions could reduce the chance of revealing the
underlying visual pattern because of uneven data points. Although the
power of visual test under the discrete uniform distribution is the
lowest in the case of non-linearity patterns, it has a relatively great
power this time. Considering for heteroskedasticity, the visual patterns
are usually detected by connecting the maximum and minimum residuals
separately at different fitted values. It will not be greatly affected
by the discrete uniform distribution compared to the uniform
distribution. Visual tests have low power on lineups with the normal
distribution and the lognormal distribution. This suggests the use of
transformation of fitted values in reading residual plots. In the case
of conventional tests, the discrete uniform distribution consistently
perform well, followed by the uniform distribution, the normal
distribution and the lognormal distribution.

In regards of the type of the simulation model, the power of visual
tests under different fitted value distributions are always lower than
the corresponding conventional tests, which is consistent with the
result reported in the previous sections.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-x-dist-poly-power-1}

}

\caption{Comparison of power on lineups with different fitted value distributions, for conventional tests and visual tests. Each power curve of the uniform distribution is surrounded by a grey line to indicate it is the baseline in its group. For both types of tests power does change. The discrete fitted value distribution substantially lowers the power of visual tests for non-linearity patterns in residual plots. For heteroskedasticity patterns, normal and lognormal distributions produce lower power for both types of tests.}\label{fig:different-x-dist-poly-power}
\end{figure}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  summarise paper content
\end{enumerate}

This paper has demonstrated that conventional residual-based statistical
tests are more sensitive to departures from assumptions for residuals
than visual tests evaluated by humans in the diagnostics of linear
regression models. An experiment is conducted in three data collection
periods with two simulation models designed for non-linearity and
heteroskedasticity model defects being applied. Methods for obtaining
the \(p\)-value of visual tests proposed by
\citet{vanderplas2021statistical} are extended to allowed for multiple
selections in a lineup protocol.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  briefly describe what was learned
\end{enumerate}

Throughout the three data collection periods of the human subject
experiment we conducted, residual plots rejected by visual tests are
rarely not rejected by conventional tests, but departures of residuals
that are of little practical sense to humans will often be alarmed by
conventional tests, supporting the common guidance given by experts to
utilize graphical representations of data in model diagnostics.

We have also found that residual-based conventional tests relying on a
user-parameterized test formula, such as the RESET test, will suffer
from a significant decrease in power for detecting certain residual
departures intended to be tested but not covered by the test formula.
This can not be avoided without prior knowledge of their functional
forms, which is usually obtained by plotting the residuals. On the other
hand, visual tests do not rely on such information and a collection of
departures can be tested simultaneously.

fitted value distribution

orientation of heter shape

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  how does this apply more broadly than simple regression
\end{enumerate}

recommend residual vs fitted values a. measure uncertainty b.
systemically miss some structure

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  where does this lead? Computer vision residual plot reading
\end{enumerate}

\hypertarget{acknowledgement}{%
\section{Acknowledgement}\label{acknowledgement}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  software tools
\item
  point to github repo (create a new one)
\item
  summarise the supplementary material
\end{enumerate}

\newpage

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  put \#eval in appendix
\item
  examine batch effect in appendix
\item
  compute the power of RESET test with different orders (considering
  dof) (focus on simple pattern possibly?)
\end{enumerate}

\bibliographystyle{tfcad}
\bibliography{paper.bib}

\end{document}