### 2023-03-21 Milestone_2/milestone_2_report.Rmd
milestone 2 template

---
title: "Milestone report"
output: pdf_document
date: "2023-03-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

A short report (5 - 10) pages

Advances in Artificial Intelligence for Data Visualization: Automate Reading of Data Plots

# Overview of the thesis

1. background and motivation
- the role of visual diagnostics in modelling
- a brief introduction to visual inference
- limitation of lineup protocol
2. research questions
-  a. study the use of visual inference in regression diagnostics with a comparison to conventional test
-  b. build a computer vision system to evaluate residual plot
-  c. make the system available to the public
3. summarise the work in the first year
- an experiment to compare visual test with conventional test 
- a prototype model for lineup evaluation
- points out the limitation of the work
4. summarise the work in the second year
- refine and reconduct the experiment
5. main findings to date
- findings from the first year
- findings from the second year

# Thesis structure

topic

chapter 1: Introduction

chapter 2
the first paper

chapter 3
build a computer vision system to evaluate residual plot

chapter 4?
software publication

chapter 5
discussion and future work

# Timetable

- April: submit abstract to ASC
- May: submit paper
- June: submit poster/short paper to IEEE vis conf, start working on computer vision model
- July: leave for few weeks
- Aug
- Sep: finalize the computer vision model
- Oct: IEEE vis conf
- Nov: web interface development 
- Dec: ASC

- Jan, 2024: 
- Feb, 2024: 
- Mar, 2024: submit paper
- Aug, 2024: submit thesis

# Difficulties

things that are actually quite serious

1. identified difficulties
2. suggestions for overcoming these difficulties