### 2024-02-21 Milestone_presentation_3/index.qmd
finish p3

---
format: 
  revealjs:
    transition-speed: fast
    slide-number: c/t
    css: custom.css
    transition: fade
    incremental: true 
    theme: default
    footer: "Slides URL: <https://patrickli-milestone-3.netlify.app>"
    logo: figures/monash-stacked-blue-rgb.png
    self-contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE
)

library(tidyverse)
```

## Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision {.center style="text-align: right;"}

::: {style="color: #5A5A5A"}

Pre-submission Review

:::

#### Weihao (Patrick) Li {style="margin-top: 40px; font-size: 0.9em"}

::: {style="font-size: 0.9em"}

Monash University

:::

---

## ✍️Supervisors

::: {.columns style="font-size:50%"}

::: {.column width="25%"}

![](figures/dicook.jpg){style="object-fit: cover; width: 100%; aspect-ratio: 1 / 1;"}

Professor Dianne Cook, Department of Econometrics and Business Statistics, Melbourne, Monash University, Australia

:::

::: {.column width="25%"}

![](figures/emitanaka.jpg){style="object-fit: cover; width: 100%; aspect-ratio: 1 / 1;"}

Dr. Emi Tanaka, Biological Data Science Institute, Australian National University, Canberra, Australia

:::

::: {.column width="25%"}

![](figures/susan.jpg){style="object-fit: cover; width: 100%; aspect-ratio: 1 / 1;"}

Assistant Professor Susan VanderPlas, Statistics Department, University of Nebraska, Lincoln, USA

:::

::: {.column width="25%"}

![](figures/klaus.jpeg){style="object-fit: cover; width: 100%; aspect-ratio: 1 / 1;"}

Senior Lecturer Klaus Ackermann, Department of Econometrics and Business Statistics, Melbourne, Monash University, Australia

:::

:::

---

## 📚Thesis Structure

1. Exploring the application of visual inference in regression diagnostics and comparing it with conventional hypothesis tests.

2. **Designing an automated visual inference system to assess residual plots of classical normal linear regression model.**

3. Deploying the automatic visual inference system as an online application and publishing the relevant open-source software.

::: {.fragment}

This presentation will be focused on the **second project**.

:::

---

## 🔍Regression Diagnostics

Diagnostics are the key to determining whether there is anything **importantly wrong** with a regression model.

<br>

$$\underbrace{\boldsymbol{e}}_\textrm{Residuals} = \underbrace{\boldsymbol{y}}_\textrm{Observations} - \underbrace{f(\boldsymbol{x})}_\textrm{Fitted values}$$

**Graphical approaches (plots)** are the recommended methods for diagnosing residuals.

---

## 🤔Challenges

:::: {.columns}

::: {.column width="40%"}

```{r fig.width=5, fig.height=5, fig.retina=3, warning=FALSE, message=FALSE}
library(tidyverse)
library(visage)
set.seed(452)
ori_x <- rand_lognormal()
mod <- heter_model(b = 0, x = closed_form(~-ori_x))
ori_dat <- mod$gen(300)

ori_dat %>%
  VI_MODEL$plot(theme = theme_light(base_size = 18), size = 1, remove_grid_line = TRUE, ) +
  # geom_line(aes(x = .fitted, y = (3.5 + 0.3 * .fitted)), col = "red") +
  # geom_line(aes(x = .fitted, y = -(3.5 + 0.3 * .fitted)), col = "red") +
  xlab("Fitted values") +
  ylab("Residuals")
```

:::

::: {.column width="60%"}

- Vertical spread of the points varies with the fitted values indicates **the existence of heteroskedasticity**.

- However, this is an **over-interpretation**.

- The visual pattern is caused by a **skewed distribution of the predictor**.

:::

::::

---

## 🔬Visual Inference

::: {style="font-size:70%"}

The reading of residual plots can be **calibrated** by an **inferential framework** called **visual inference** (Buja, et al. 2009).

:::: {.columns}

::: {.column width="50%"}

```{r fig.width=5, fig.height=5, fig.retina=3}
set.seed(452)
mod$gen_lineup(300, k = 20, pos = 11) %>%
  filter(null != FALSE) %>%
  bind_rows(ori_dat %>% mutate(k = 11, null = FALSE)) %>%
  VI_MODEL$plot_lineup(theme = theme_light(base_size = 12),
                       remove_grid_line = TRUE,
                       remove_axis = TRUE)
```

:::

::: {.column width="50%"}

::: {.fragment}

Typically, a **lineup** of residual plots consists of

<!-- - $m$ randomly placed plots -->
- one **data plot**
- $19$ **null plots** containing residuals **simulated from the fitted model**.

:::

::: {.fragment}

To perform a visual test

- Observer(s) will be asked to select the **most different plot(s)**.
- The p-value can be calculated using the **beta-binomial model** (VanderPlas et al., 2021).

:::

:::

::::

:::

---

## 🚫Limitations of Lineup Protocol

::: {.fragment}
1. Human can not

- evaluate lineup consisted of a **large number of plots**.
- evaluate a **large number of lineups**.
:::

::: {.fragment}
2. Evaluation of lineup is **high in labour cost** and **time consuming**.
:::

---

## 🤖Computer Vision Model

Modern **computer vision models** are well-suited for addressing this challenge.

It is usually built on a **deep neural network** called **convolutional neural network** (CNN).

![Source: https://en.wikipedia.org/wiki/Convolutional_neural_network](figures/Typical_cnn.png)

---

## 📏Measure the Difference

To develop a computer vision model for assessing lineups of residual plots, we need to define a **numerical measure** of **"difference"** or **"distance"** between plots.

- pixel-wise sum of square differences
- Structural Similarity Index Measure (SSIM)
- scagnostics
- metric learning
- ...

---

## 🎲Residual Distribution

Consider the **classical normal linear regression model**

$$\boldsymbol{y} = \boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon}, \quad \boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n).$$

By the **Frisch-Waugh-Lowell theorem**,

$$\boldsymbol{e} = \boldsymbol{R}\boldsymbol{\varepsilon},$$
where $\boldsymbol{R}=\boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$.

We treat $\boldsymbol{e}$ as **a vector of random variables** here.

---

## 🎲Residual Distribution

::: {style="font-size:80%"}

The **minimization of the sum of square residuals** implies  
$$\sum_{i=1}^{n} e_i = 0 \quad\text{and}\quad \text{rank}(\boldsymbol{R}) = n - 1.$$

Thus, $\boldsymbol{e}$ follows a **degenerate multivariate distribution**.

For simplicity, we will replace $\text{cov}(\boldsymbol{e}, \boldsymbol{e}) = \boldsymbol{R}\sigma^2\boldsymbol{R}' = \boldsymbol{R}\sigma^2$ with a **full-rank diagonal matrix** $\text{diag}(\boldsymbol{R}\sigma^2)$.

Then for a **correctly specified model**, $$\boldsymbol{e} \sim N(\boldsymbol{0}_n, \text{diag}(\boldsymbol{R}\sigma^2)).$$

Symbol $Q$ will be used to represent this **reference residual distribution**.

:::

---

## 🎲Residual Distribution

However, if the model is **misspecified**, then the **actual residual distribution** denoted as $P$, will be **different** from $Q$.

- For example, if $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\boldsymbol{V})$, where $\boldsymbol{V} \neq \sigma^2\boldsymbol{I}_n$,  $$\boldsymbol{e} \sim N(\boldsymbol{0}_n, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R})) \implies \text{Heteroskedasticity}.$$

- And if some necessary higher-order predictors $\boldsymbol{Z}$ are also omitted, $$\boldsymbol{e} \sim N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R})) \implies \text{Non-linearity}~ \& ~\text{Heteroskedasticity}.$$

---

## 📏KL Divergence of $P$ from $Q$

We defined a **distance measure** based on **Kullback-Leibler divergence** to quantify the **extent of model violations**

\begin{align}
\label{eq:kl-0}
D &= \log\left(1 + D_{KL}\right), \\
\label{eq:kl-1}
D_{KL} &= \int_{\mathbb{R}^{n}}\log\frac{p(\boldsymbol{e})}{q(\boldsymbol{e})}p(\boldsymbol{e})d\boldsymbol{e},
\end{align}

\noindent where $p(.)$ and $q(.)$ are the probability density functions for distribution $P$ and $Q$ respectively.

- $D = 0$ if and only if $P \equiv Q$.

---

## 📝Evaluation of $D$ for Normal $\boldsymbol{\varepsilon}$

::: {style="font-size:80%"}

For a **classical normal linear regression model** that omits necessary higher-order predictors $\boldsymbol{Z}$, and incorrectly assumes $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$ while in fact $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})$ with $\boldsymbol{V} \neq \sigma^2\boldsymbol{I}_n$, the KL divergence can be further expanded to

\begin{align}
\label{eq:kl-2}
D_{KL} &= \frac{1}{2}\left(\log\frac{|\boldsymbol{W}|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\boldsymbol{W}^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'\boldsymbol{W}^{-1}\boldsymbol{\mu}_z\right),
\end{align}

where $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$, and $\boldsymbol{W} = \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R})$.

- The assumed error variance $\sigma^2$ is set to be $\text{tr}(\boldsymbol{V})/n$, which is the expectation of the estimated variance.

:::

---

## 📝Evaluation of $D$ for Non-normal $\boldsymbol{\varepsilon}$

For non-normal $\boldsymbol{\varepsilon}$, $P$ is **unlikely to be a well-known distribution**.

- difficult to compute $p(\boldsymbol{e})$. 
- difficult to do numerical integration over the $n$ dimensional space, because $n$ could be **potentially very large**.

::: {.fragment}

Therefore, the elements of $\boldsymbol{e}$ are assumed to be **independent of each other**.

:::

---

## 📝Evaluation of $D$ for Non-normal $\boldsymbol{\varepsilon}$

::: {style="font-size:80%"}

Under the independence assumption, the **integration becomes a summation**

\begin{align}
\label{eq:kl-3}
D_{KL} &\approx \sum_{i = 1}^{n} \hat{D_{KL}^{(i)}}, \\
\hat{D_{KL}^{(i)}} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
\end{align}

where $B$ is a matrix with $n$ rows and $m$ columns, and **each column is a set of simulated residuals**, $m$ is the **number of simulations**, $\hat{p_i}(.)$ is the **kernel density estimator** of $p_i(.)$ using simulated residuals in $B_i$.

- The assumed variance for $q(.)$ is set to be $\sum_{b \in vec(B)}(b - \sum_{b \in vec(B)} b/nm)^2/(nm - 1)$.

:::

---

## 🎯Approximation of the Distance

::: {style="font-size:80%"}

Let's get back to the **original problem**. We have a fitted model on hand, and we want to know **how different the residuals are from a set of good residuals**.

- **However, the data generating process is typically unknown!**

  - The distance can not be computed!

::: {.fragment}

We can train a **computer vision model** to approximate $D$ with **a residual plot**

\begin{equation}
\label{eq:d-approx}
\widehat{D} = f_{CV}(V_{h \times w}(\boldsymbol{e}, \boldsymbol{\hat{y}})),
\end{equation}

where $V_{h \times w}(.)$ is a **plotting function** that saves a residual plot as an image with $h \times w$ pixels, and  $f_{CV}(.)$ is a **computer vision model** which predicts distance in $[0, +\infty)$.

:::

:::

---

## 🔄Generation of Training Data

::: {style="font-size:80%"}

### **Data generating process**

\begin{align}
\label{eq:data-sim}
\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 + \beta_2(\boldsymbol{z} + \beta_1\boldsymbol{w}) + \boldsymbol{k} \odot \boldsymbol{\varepsilon}, \\
\boldsymbol{z} &\propto \text{He}_j(\boldsymbol{x}_1), \\
\boldsymbol{w} &\propto \text{He}_j(\boldsymbol{x}_2), \\
\boldsymbol{k} &= \sqrt{\boldsymbol{1}_n + b(2 - |a|)(\boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 - a\boldsymbol{1}_n)^2},
\end{align}

where $\text{He}_j(.)$ is the $j$th-order **probabilist's Hermite polynomials**, and the $\sqrt{(.)}$ and $(.)^2$ operators are element-wise operators.

- The residuals are obtained by **regressing $y$ on $x_1$**. If $\beta_1 \neq 0$, $x_2$ will also be included as a predictor.

:::

---

## 💡Non-linearity

:::: {.columns}

::: {.column width="30%"}

**Factor $j$**

:::

::: {.column width="70%"}

```{r fig.width=10, fig.height=4}
set.seed(10086)

# Data for shape 1
dat_shape_1 <- phn_model(j = 2, include_x2 = FALSE, sigma = 0.05)$gen(500) %>%
  mutate(j = 2)

# Generate data for shape 2, 3 and 4. Reuse x and e.
map_df(3:11, function(j) {
  phn_model(j = j, include_x2 = FALSE, sigma = 0.05)$
    gen(500, computed = select(dat_shape_1, x1, e)) %>%
  mutate(j = j)
}) %>%

# Combined with data for shape 1
bind_rows(dat_shape_1) %>%
  mutate(j = factor(j)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light(base_size = 15)) +
  facet_wrap(~j, scales = "free", labeller = label_parsed, ncol = 5)
```

:::

:::

:::: {.columns}

::: {.column width="30%"}

**Factor $\sigma_\varepsilon$**

:::

::: {.column width="70%"}

```{r fig.width=8, fig.height=2}
set.seed(10086)

# Data for shape 1
dat_shape_1 <- phn_model(j = 2, include_x2 = FALSE, sigma = 0.05)$gen(500) %>%
  mutate(e_sigma = 0.2)

# Generate data for shape 2, 3 and 4. Reuse x and e.
map_df(c(0.4, 0.8, 1.6, 3.2), function(e_sigma) {
  phn_model(j = 2, include_x2 = FALSE, sigma = e_sigma)$
    gen(500, computed = select(dat_shape_1, x1, e) %>% mutate(e = e/0.2*e_sigma)) %>%
  mutate(e_sigma = e_sigma)
}) %>%

# Combined with data for shape 1
bind_rows(dat_shape_1) %>%
  mutate(e_sigma = factor(e_sigma)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light(base_size = 15)) +
  facet_wrap(~e_sigma, scales = "free", labeller = label_parsed, ncol = 5)
```

:::

:::

---

## 💡Heteroskedasticity

:::: {.columns}

::: {.column width="30%"}

**Factor $a$**

:::

::: {.column width="70%"}

```{r fig.width=10, fig.height=4}
set.seed(10085)

# Generate data for a = -1
dat_a_n1 <- phn_model(include_z = FALSE,
                      include_x2 = FALSE,
                      a = -1,
                      b = 100)$gen(500) %>%
  mutate(a = -1)

# Generate data for other a
map(c(-0.75, -0.5, -0.25, 0, 0.25, 0.5, 0.75, 1), function(a) {
  phn_model(include_z = FALSE,
            include_x2 = FALSE,
            a = a,
            b = 100)$gen(500) %>%
  mutate(a = a)
}) %>%

  # Combined with data for a = -1
  bind_rows(dat_a_n1) %>%
  mutate(a = factor(a)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light(base_size = 15)) +
  facet_wrap(~a, scales = "free", ncol = 5) +
  xlab("Fitted values") +
  ylab("Residuals")
```

:::

:::

:::: {.columns}

::: {.column width="30%"}

**Factor $b$**

:::

::: {.column width="70%"}

```{r fig.width=8, fig.height=2}
set.seed(10085)

# Generate data for a = -1
dat_a_n1 <- phn_model(include_z = FALSE,
                      include_x2 = FALSE,
                      a = -1,
                      b = 6)$gen(500) %>%
  mutate(b = 6)

# Generate data for other a
map(c(3, 1.5, 1, 0.5), function(b) {
  phn_model(include_z = FALSE,
            include_x2 = FALSE,
            a = -1,
            b = b)$gen(500) %>%
  mutate(b = b)
}) %>%

  # Combined with data for a = -1
  bind_rows(dat_a_n1) %>%
  mutate(b = factor(b)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light(base_size = 15)) +
  facet_wrap(~b, scales = "free", ncol = 5) +
  xlab("Fitted values") +
  ylab("Residuals")
```

:::

:::

---

## 💡Non-normality

:::: {.columns}

::: {.column width="30%"}

**Distribution of $\varepsilon$**

:::

::: {.column width="70%"}

```{r fig.width=8, fig.height=6}
set.seed(10086)

# Data for shape 1
dat_shape_1 <- phn_model(include_z = FALSE, include_x2 = FALSE, e = rand_uniform(-1.4, 1.4))$gen(500) %>%
  mutate(e_dist = "uniform")

dat_shape_2 <- phn_model(include_z = FALSE, include_x2 = FALSE, sigma = 0.8)$gen(500) %>%
  mutate(e_dist = "normal")

dat_shape_3 <- phn_model(include_z = FALSE, include_x2 = FALSE, e = rand_lognormal(sigma = 0.6))$gen(500) %>%
  mutate(e_dist = "lognormal")

dat_shape_4 <- phn_model(include_z = FALSE, include_x2 = FALSE, e = rand_uniform_d(-1.4, 1.4, even = TRUE))$gen(500) %>%
  mutate(e_dist = "discrete")

# Generate data for shape 2, 3 and 4. Reuse x and e.
bind_rows(dat_shape_1, dat_shape_2, dat_shape_3, dat_shape_4) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light(base_size = 15)) +
  facet_wrap(~e_dist, scales = "free", labeller = label_parsed, ncol = 2)
```

:::

:::

---

## 💡Multiple Model Violations

::: {.columns}

::: {.column width="30%"}

**Non-linearity + Heteroskedasticity**

:::

::: {.column width="70%"}

```{r fig.width=10, fig.height=4}
set.seed(10086)

# Data for shape 1
dat_shape_1 <- phn_model(j = 2, a = -1, b = 100, include_x2 = FALSE, sigma = 0.05)$gen(500) %>%
  mutate(j = 2)

# Generate data for shape 2, 3 and 4. Reuse x and e.
map_df(3:11, function(j) {
  phn_model(j = j, a = -1, b = 100, include_x2 = FALSE, sigma = 0.05)$
    gen(500, computed = select(dat_shape_1, x1, e)) %>%
  mutate(j = j)
}) %>%

  # Combined with data for shape 1
  bind_rows(dat_shape_1) %>%
  mutate(j = factor(j)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light(base_size = 15)) +
  facet_wrap(~j, scales = "free", labeller = label_parsed, ncol = 5)
```

:::

:::

::: {.columns}

::: {.column width="30%"}

**Non-normality + Heteroskedasticity**

:::

::: {.column width="70%"}

```{r fig.width=8, fig.height=2}
set.seed(10085)

# Data for shape 1
dat_shape_1 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, e = rand_uniform(-1.4, 1.4))$gen(500) %>%
  mutate(e_dist = "uniform")

dat_shape_2 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, sigma = 0.8)$gen(500) %>%
  mutate(e_dist = "normal")

dat_shape_3 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, e = rand_lognormal(sigma = 0.6))$gen(500) %>%
  mutate(e_dist = "lognormal")

dat_shape_4 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, e = rand_uniform_d(-1.4, 1.4, even = TRUE))$gen(500) %>%
  mutate(e_dist = "discrete")

# Generate data for shape 2, 3 and 4. Reuse x and e.
bind_rows(dat_shape_1, dat_shape_2, dat_shape_3, dat_shape_4) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light(base_size = 15)) +
  facet_wrap(~e_dist, scales = "free", labeller = label_parsed, ncol = 4)
```

:::

:::

---

## 💡Predictor Distribution

::: {.columns}

::: {.column width="30%"}

**Distribution of predictor**

:::

::: {.column width="70%"}

```{r fig.width=8, fig.height=6}
set.seed(10086)

stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

# Data for uniform distribution
dat_dist_1 <- poly_model(shape = 1, 
                      x = {
                        raw_x <- rand_uniform(-1, 1);
                        closed_form(~stand_dist(raw_x))
                        }, 
                      sigma = 0.5)$gen(300) %>%
  mutate(x_dist = "uniform")

# Generate data for other distributions
dat_dist_2 <- poly_model(shape = 1,
                         x = {
                           raw_x <- rand_normal(sigma = 0.3); 
                           closed_form(~stand_dist(raw_x))
                           }, 
                         sigma = 0.5)$
  gen(300, computed = select(dat_dist_1, e)) %>%
  mutate(x_dist = "normal")

dat_dist_3 <- poly_model(shape = 1,
                         x = {
                           raw_x <- rand_lognormal(sigma = 0.6); 
                           closed_form(~stand_dist(raw_x/3 - 1))
                           }, 
                         sigma = 0.5)$
  gen(300, computed = select(dat_dist_1, e)) %>%
  mutate(x_dist = "lognormal")

dat_dist_4 <- poly_model(shape = 1,
                         x = {
                           raw_x <- rand_uniform_d(k = 5, even = TRUE); 
                           closed_form(~stand_dist(raw_x))
                           }, 
                         sigma = 0.5)$
  gen(300, computed = select(dat_dist_1, e)) %>%
  mutate(x_dist = "discrete")

# Generate and plot data for discrete uniform distribution
bind_rows(dat_dist_1, dat_dist_2, dat_dist_3, dat_dist_4) %>%
  mutate(x_dist = factor(x_dist, 
                         levels = c("discrete", 
                                    "lognormal", 
                                    "normal", 
                                    "uniform"))) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light(base_size = 15)) +
  xlab("Fitted values") +
  ylab("Residuals") +
  facet_wrap(~x_dist, ncol = 2, scales = "free")
```

:::

:::

---

## 💡Second Predictor

::: {.columns}

::: {.column width="30%"}

**Non-linearity + Second predictor**

:::

::: {.column width="70%"}

```{r fig.width=10, fig.height=4}
set.seed(10086)

# Data for shape 1
dat_shape_1 <- phn_model(j = 2, include_x2 = TRUE, sigma = 0.05)$gen(500) %>%
  mutate(j = 2)

# Generate data for shape 2, 3 and 4. Reuse x and e.
map_df(3:11, function(j) {
  phn_model(j = j, include_x2 = TRUE, sigma = 0.05)$
    gen(500, computed = select(dat_shape_1, x1, e)) %>%
  mutate(j = j)
}) %>%

  # Combined with data for shape 1
  bind_rows(dat_shape_1) %>%
  mutate(j = factor(j)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light(base_size = 15)) +
  facet_wrap(~j, scales = "free", labeller = label_parsed, ncol = 5)
```

:::

:::

---

## 🏛️Model Architecture

The architecture of the computer vision model is adapted from **VGG16** (Simonyan and Zisserman 2014).

```{r fig.align='center'}
magick::image_read_pdf("figures/cvm.pdf", pages = 1)
```

---

## 🏋️‍♂️Model Training

The computer vision model is trained on the **M3 high-performance computing platform** (www.massive.org.au), using **TensorFlow** (Abadi et al. 2016) and **Keras** (Chollet et al. 2015).

- The training, validation and test set contains 64000, 16000 and 8000 images respectively.

- The distribution of the target variable $D$ is controlled such that it roughly follows a uniform distribution.

- Multiple models with different image resolutions are trained, the optimized model has input size $32 \times 32$.

---

## 🌐Overview of Model Performance

::: {.columns}

::: {.column width="40%"}

::: {style="font-size:80%"}

The model performs

- better for non-normality
- worse for non-linearity
- consistently across different datasets

::: {.fragment}

The model performs worst on null plots (D = 0) as expected.

:::

:::

:::

::: {.column width="60%"}

```{r}
library(glue)
test_pred <- readr::read_csv(here::here("data/test_pred.csv"))
train_pred <- readr::read_csv(here::here("data/train_pred.csv"))
meta <- readr::read_csv(here::here("data/meta.csv"))

model_pred <- train_pred %>% 
  left_join(meta) %>%
  mutate(type = "train") %>%
  bind_rows(test_pred %>% 
              left_join(meta) %>%
              mutate(type = "test")) %>%
  filter(res == 32L)

model_pred %>%
  group_by(type, include_non_normal, include_heter, include_z) %>%
  summarise(n = n(), 
            RMSE = yardstick::rmse_vec(effect_size, vss)) %>%
  mutate(violations = ifelse(include_z, "non-linearity", "null")) %>%
  mutate(violations = ifelse(include_heter, glue("{violations} + heteroskedasticity"), violations)) %>%
  mutate(violations = ifelse(include_non_normal, glue("{violations} + non-normality"), violations)) %>%
  mutate(violations = gsub("null \\+ ", "", violations)) -> model_pred_plot_dat

```

```{r fig.height=6, fig.width=7.2}

grey_grid <- expand.grid(c(0,1,2,3,4,5,6), c(1,2,3))

model_pred_plot_dat <- model_pred_plot_dat %>%
  mutate(violations = factor(violations, levels = c("null", 
                                                    "non-linearity",
                                                    "heteroskedasticity",
                                                    "non-normality",
                                                    "non-linearity + heteroskedasticity",
                                                    "non-linearity + non-normality",
                                                    "heteroskedasticity + non-normality",
                                                    "non-linearity + heteroskedasticity + non-normality")))

ggplot() +
  geom_ribbon(aes(x = seq(-0.5, 7.5, 0.5), ymin = 0.5, ymax = 1.5), fill = "grey90") +
  geom_ribbon(aes(x = seq(-0.5, 7.5, 0.5), ymin = 2.5, ymax = 3.5), fill = "grey90") +
  geom_point(aes(grey_grid$Var1, grey_grid$Var2), col = "grey", size = 5) +
  geom_point(aes(c(1,2,3), c(1,2,3)), size = 5) +
  geom_point(aes(c(4,4,5,5,6,6,7,7,7), c(1,2,1,3,2,3,1,2,3)), size = 5) +
  geom_segment(aes(x = c(4,5,6,7,7), xend = c(4,5,6,7,7), y = c(1,1,2,1,2), yend = c(2,3,3,2,3)), size = 1) +
  geom_text(aes(c(-2, -2, -2), c(1, 2, 3), label = c("Non-linearity", "Heteroskedasticity", "Non-normality"))) +
  geom_segment(aes(x = -0.5, xend = 7.5, y = 3.5, yend = 3.5)) +
  geom_segment(aes(x = -0.5, xend = -0.5, y = 3.5, yend = 3.5 + 1.5 * 5)) +
  geom_segment(aes(x = rep(-0.6, 4), xend = rep(-0.5, 4), y = 3.5 + seq(0, 1.5, 0.5) * 5, yend = 3.5 + seq(0, 1.5, 0.5) * 5)) +
  geom_text(aes(x = rep(-1, 4), y = 3.5 + seq(0, 1.5, 0.5) * 5, label = c("  0", "0.5", "  1", "1.5")), size = 4) +
  geom_point(data = model_pred_plot_dat, aes(as.numeric(violations) - 1, RMSE * 5 + 3.5, col = type), size = 3) +
  geom_text(aes(-1.5, 0.75 * 5 + 3.5, label = "RMSE"), angle = 90) +
  theme_light(base_size = 15) +
  theme(panel.grid = element_blank(), 
        axis.text = element_blank(), 
        axis.ticks = element_blank()) +
  xlab("") +
  ylab("") +
  xlim(c(-4, 8)) +
  labs(col = "Dataset")
```

:::

:::

---

## 📈Model Violations Index

::: {style="font-size:80%"}

For a consistent data generating process, $D$ typically **increases logarithmically with the number of observations**.

This behaviour comes from the relationship $D = \text{log}(1 + D_{KL})$, where $D_{KL} = \sum_{i=1}^{n}D_{KL}^{(i)}$ under the assumption of independence.

Therefore, the **Model Violations Index** (MVI) for determining the **extent of model violations** can be proposed as

\begin{equation}
\label{eq:mvi}
\text{MVI} = C + \widehat{D} - log(n),
\end{equation}

where $C$ is a large enough constant keeping the result positive.

:::

---

## 📈Model Violations Index

```{r fig.align='center'}
magick::image_read_pdf("figures/poly-index-1.pdf", pages = 1)
```

---

## 📈Model Violations Index

```{r fig.align='center'}
magick::image_read_pdf("figures/heter-index-1.pdf", pages = 1)
```

---

## 🔬Statistical Testing Based on the Approximated Distance

::: {style="font-size:80%"}

We have proposed a statistical test based on the approximated distance

- The **null distribution** can be estimated by predicting approximated distance for **a large number of null plots**.

- The **critical value** can be estimated by the **sample quantile** (e.g. $Q_{null}(0.95)$) of the null distribution.

- The **$p$-value** is the proportion of null plots having approximated distance **greater than or equal to** the observed one.

:::

---

## 🔬Statistical Testing Based on the Approximated Distance

::: {.columns}

::: {.column width="100%"}

```{r}
my_vi <- autovi::auto_vi(lm(mpg ~ ., data = mtcars))
my_vi$check_result <- readRDS("data/test.rds")

p1 <- my_vi$plot_resid()
p2 <- my_vi$summary_plot() +
  ggtitle("", subtitle = glue("P-value = {my_vi$check_result$p_value}")) +
  theme(legend.position = "bottom", legend.box = "vertical") +
  labs(linetype = "") +
  xlab(expression(hat(D))) +
  scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D))))

patchwork::wrap_plots(p1, p2)
```

:::

:::

---

## 🧪Human Subject Experiment

We have collected $7974$ responses to $1152$ lineups in a **human subject experiment** conducted in the first project (Li et al. 2023).

Differences:

- non-linearity and heteroskedasticity **do not co-exist in one lineup**.
- non-normality and multiple predictors **are not considered** in the experimental design.

---

## 🧪Compared to Visual Tests

```{r fig.align='center'}
magick::image_read_pdf("figures/human-mosaic-1.pdf", pages = 1)
```

---

## 🧪Compared to RESET and BP Tests

```{r fig.align='center'}
magick::image_read_pdf("figures/conv-mosaic-1.pdf", pages = 1)
```

---

## 🧪Power Comparison

```{r fig.align='center'}
magick::image_read_pdf("figures/power-1.pdf", pages = 1)
```

---

## 💡Example: Left-triangle

```{r fig.align='center'}
magick::image_read_pdf("figures/false-check-1.pdf", pages = 1)
```

## 💡Example: Left-triangle

```{r fig.align='center'}
magick::image_read_pdf("figures/false-lineup-1.pdf", pages = 1)
```

## 💡Example: Boston Housing

$MEDV = \beta_0 + \beta_1RM + \beta_2LSTAT + \beta_3PTRATIO + \varepsilon$

```{r fig.align='center'}
magick::image_read_pdf("figures/boston-check-1.pdf", pages = 1)
```

## 💡Example: Boston Housing

```{r fig.align='center'}
magick::image_read_pdf("figures/boston-lineup-1.pdf", pages = 1)
```

## 💡Example: Dinosaur

$y = \beta_0 + \beta_1x + \varepsilon$

```{r fig.align='center'}
magick::image_read_pdf("figures/dino-check-1.pdf", pages = 1)
```

## 💡Example: Dinosaur

```{r fig.align='center'}
magick::image_read_pdf("figures/dino-lineup-1.pdf", pages = 1)
```

---

## 🎬Conclusions

::: {style="font-size:80%"}

In this study, we have

- introduced a **distance measure** to effectively **captures the magnitude of model violations**

- proposed and trained a **computer vision model** to **approximate this distance**

- developed a **statistical testing procedure** and a **Model Violations Index** based on the approximated distance

:::

---

## 🎬Conclusions

::: {style="font-size:80%"}

- The proposed test shows **comparable power** to conventional ones, but there's room for further **enhancing its robustness** against **minor deviations** from model assumptions.

- Our method offers significant value by **easing analysts' workload** in **assessing residual plots**.

- While we encourage **continued manual review of residual plots** for **insightful analysis**, our approach serves as a valuable tool for **automating** or **supplementing** the diagnostic process.

:::

---

## 📅Timeline

```{r}
data.frame(Date = c("Mar",
                    "April", 
                    "May",
                    "July",
                    "",
                    "Aug"), 
           Event = c("Finish and submit the second paper",
                     "Develop a web interface for the automatic visual inference system",
                     "Clean and publish R packages to CRAN",
                     "Finish the third paper",
                     "Attend the useR! conference",
                     "Submit thesis")) %>%
  kableExtra::kable(booktabs = TRUE) %>%
  kableExtra::kable_styling() 
```

## Thanks! Any questions? {.center}

<br>

### 🔗Relevant links

::: {.nonincremental}

- 📦 R Packages
  - [`visage`: Visual Inference for Linear Regression Diagnostics](https://github.com/TengMCing/visage){target="_blank"}
  - [`autovi`: Automated Visual Inference with Computer Vision Models](https://github.com/TengMCing/autovi){target="_blank"}

- ![](figures/github-mark/github-mark.png){height=1em width=1em style="vertical-align:middle"} [Slides](https://github.com/TengMCing/PhD/Milestone_presentation_3){target="_blank"}

:::