### 2024-08-22 Thesis/05-chap5.qmd
skeleton of conclusion chapter

# Conclusion and future plans

The three pieces of work in this thesis focus on ...

## Contributions

The primary contributions of this research is ....

Principles of transparency and reproducible research have guided the work. The materials to reproduce this thesis are available at .... and software packages are:

- software 1
- software 2

...

## Future work

There are several directions that this work can be developed: alternative  computer vision models, and extensions to different types of visual diagnostics.

The limitations of the current are ... We have explored these other ways to make the model ... Some new possibilities are ... might address ...

Residual diagnostics are also used in more complex statistical models such as hierarchical, or temporal or spatial ... These require adaptations of residual calculations ...

Visual diagnostics are also common in Bayesian modeling to assess ... (eg Gelman, A. 2004 Exploratory data analysis for complex models. J. Comput. Graph. Stat. 13(4), 755–779.)