### 2024-11-28 Thesis/amend.qmd
add amend summary

---
title: "Amendments Summary"
author: Weihao Li
Date: "`r Sys.Date()`"
format: pdf
---

In this version, I have added a paragraph to acknowledge the extensive literature on regression diagnostics and various types of regression models. I clarified that the thesis focuses specifically on the classical normal linear regression model.

I also improved the language in Chapter 3 for better clarity, such as changing "dynamic input size" to "variable input size." Additionally, I included a sentence to provide more details on how the predictors in the training datasets are generated.

In Chapter 4, I corrected several formatting issues, particularly those related to code chunks not displaying properly.

Overall, the reviewers' comments have been very helpful, though not all could be addressed due to time constraints. However, as we are preparing the journal submission for Chapters 3 and 4, these comments will be carefully considered and properly addressed.